---
title: 'T-Cell Fluorescence Oscillation Analysis'
author: "Jing Zhao"
date: "10/20/2018"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(
  timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste('Time for this code chunk:', as.character(res))
    }
  }})
)

library(Rssa)
library(ggplot2)
library(dplyr)
library(reshape)
library("scales")
library(gridExtra)
library(DescTools)
library("EBImage")

```

## Set analysis parameters 
```{r, timeit = TRUE}
# Set ssa window length
window = 124
# Set lag for autocorrelation
lag = 120
# Threshold R-square value to screen sinusoid curve
threshold = 0.98
# Exposure Time (Frame rate, per ms)
frame_rate = 800
# Period Constraints (Frame rate)
low_pf = 8
high_pf = 180
```

## Define Functions
```{r,timeit = TRUE}
# Function: Extract x,y coordiantes from each Header
row.xy <- function(string){
  xx <- strsplit(string,"[.]")[[1]][1]
  yy <- strsplit(string,"[.]")[[1]][2]
  x <- as.numeric(substring(xx,2))
  y <- as.numeric(substring(yy,2))
  return(c(x,y))
}

# Function: Split the rowname column into two columns
split.rowname.xy <- function(df) {
  rowNames <- rownames(df)
  df.xy <- data.frame(t(data.frame(lapply(rowNames, row.xy))))
  names(df.xy) <- c("X","Y")
  df[["X"]] <- df.xy[["X"]]
  df[["Y"]] <- df.xy[["Y"]]
  return(df)
}

# Function: Linear detrend 
lm.detrend <- function(df) {
  names(df) <- "Intensity"
  tt <- seq(1,dim(df)[1],1) * frame_rate/1000
  df[["Time"]] <- tt
  lm.fit <- lm(Intensity ~ Time, df)
  df[["Intensity"]] <- df[["Intensity"]] - predict(lm.fit)
  return(df)
}

# Function: Pixel-wise screening for sinusoid oscillation
sine.screen <- function(ts, per, threshold) {
  
  # Conducting Fast Fourier Transform (FFT)  
  ssp <- spectrum(ts, plot=FALSE) 
  # Calculate period from FFT
  per_fft <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
  
  if (per < 24) {
    per <- per_fft
  }
  
  # The Period screening criteria !!!
  if ((per < high_pf) & (per > low_pf)) {
      acf <- acf(ts, lag.max = lag, plot = FALSE)
      # Extract the acf and time pairs 
      y <- acf$acf[1:lag]   # acf value
      t <- acf$lag[1:lag]   # lag time
      
      # Transfomred linear fit for sinusoid time-series
      df_sine <- data.frame(cbind(y, t, sin(2*pi/per*t), cos(2*pi/per*t)))
      names(df_sine) <- c('y','t','sine', 'cos')
      reslm <- lm(y ~  sine + cos + t * sine + t * cos, data = df_sine)
      # The sinusoid screening criteria !!!
      if (summary(reslm)$r.squared >= threshold) {
        return(per)
      }
      else {
        return(0)
      } 
  }
  else {
      return(0)
  }
}

# Function: Plot time-series with single dataframe
ts_plot <- function (col, title){
  df <- data.frame(col)
  names(df) <- "Intensity"
  tt <- seq(1,dim(df)[1],1)* frame_rate/1000
  df[["Time"]] <- tt
  p1 <- ggplot(df, aes(x=Time, y=Intensity))  + geom_line() + ggtitle(title) + xlab("Time (s)")
  return(p1)
}

# Function: Pixel-wise screening for sinusoid and plotting the extracted vectors and fitting 
sine.screen.plot <- function(ts, per, threshold, header) {
  acf <- acf(ts, lag.max = lag, plot = FALSE)
  # Extract the acf and time pairs 
  y <- acf$acf[1:lag]   # acf value
  t <- acf$lag[1:lag]   # lag time
  
  # Conducting Fast Fourier Transform (FFT)  
  ssp <- spectrum(ts, plot=FALSE) 
  # Calculate period from FFT
  per_fft <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
  
  if (per < 24) {
    per <- per_fft
  }
  
  print(per)
  
  # Transfomred linear fit for damped sinusoid time-series
  df_sine <- data.frame(cbind(y, t, sin(2*pi/per*t), cos(2*pi/per*t)))
  names(df_sine) <- c('y','t','sine', 'cos')
  reslm <- lm(y ~  sine + cos + t * sine + t * cos, data = df_sine)
  
  # The sinusoid screening criteria !!! 
  if (summary(reslm)$r.squared >= threshold & (per < high_pf) & (per > low_pf)) {
    # Plot the spectral periodgram of the subset singal
    df_ssp <- data.frame(cbind(ssp$spec,round(1/ssp$freq*frame_rate/1000,digits = 2)))
    names(df_ssp) <- c('Spec','Period')
    p_spec <- ggplot(df_ssp[2:125,], aes(x=Period, y=Spec)) + geom_bar(stat="identity", color="black") + xlab("Period (s)") + ylab("Spectral Density") + ggtitle(paste(header, "Periodgram,","Period =", round(per*frame_rate/1000,digits=2)))
    
    # Plot reconstrcuted vector calculated from SSA
    amp <- round(mean(abs(ts))*pi/2,digits=2) 
    p_uu <- ts_plot(ts, paste(header, "Principal Time-series, average Amplitude =", toString(amp)))
    title <- paste(header, "ACF,", "Period =", round(per*frame_rate/1000,digits=2), ",", "R-square =", toString(round(summary(reslm)$r.squared,digits = 3)))
    
    df_sine[['y_hat']] <- data.frame(fitted(reslm))

    p_sine <-ggplot(df_sine, aes(x = t*frame_rate/1000)) + geom_point(aes(y=y), shape = 1) + geom_line(aes(y=y_hat, color = "red")) + xlab("Lag (s)") + ylab("ACF") + ggtitle(title)
    grid.arrange(p_uu,p_spec,p_sine,ncol=1)
  } 
}

amplitude_sel <- function(df_stats){
  if (df_stats[18] != 0){
      if (df_stats[18] == df_stats[4]){
      df_stats[19] <- df_stats[5]
      } 
      else if (df_stats[18] == df_stats[7]){
      df_stats[19] <- df_stats[8]
      } 
      else if (df_stats[18] == df_stats[10]){
      df_stats[19] <- df_stats[11]
      } 
      else if (df_stats[18] == df_stats[13]){
      df_stats[19] <- df_stats[14]
      } 
  } else {
      df_stats[19] <- 0
  }
}

print.df.info <- function(df1, type_name) {
  # Z dimension
  print(paste("There are", dim(df1)[1]-1,"frames in the", type_name, "imaging."))
  # X - Y dimensions
  colNames <- colnames(df1)
  start_xy <- row.xy(colNames[1])
  end_xy <- row.xy(colNames[length(colNames)])
  print(paste("The x coordinstes range from", start_xy[1], "to", end_xy[1],"."))
  print(paste("The y coordinstes range from", start_xy[2], "to", end_xy[2],"."))
  x_pixels <- end_xy[1] - start_xy[1] + 1
  y_pixels <- end_xy[2] - start_xy[2] + 1
  print(paste("The dimension of image frame is", x_pixels,"*", y_pixels,"."))
}

```

## Imaging Data Loading and Preprocessing 

```{r, warning=FALSE, fig.height = 5, fig.width = 4, fig.align ="center", timeit = TRUE, eval=FALSE}

# Load data matrix
df_ROI <- read.csv("Cell-9&10.csv")       
# Remove first two columns, Header is the pixel coordinate in "x.y"" format
df1_ROI <- df_ROI[,3:dim(df_ROI)[2]]
# Report information of cell imaging dafaframe 
print.df.info(df1_ROI, "cell (ROI + Peripheral)")
cat("\n")

# Load data matrix
df_BG <- read.csv("BG-1.csv")    
# Remove first two columns, Header is the pixel coordinate in "x.y"" format
df1_BG <- df_BG[,3:dim(df_BG)[2]]
# Report information of background imaging dafaframe 
print.df.info(df1_BG, "background (BG)")

# Stack Cell and BG dataframes together
df1_combined <- cbind(df1_ROI, df1_BG)
df2 <- data.frame(t(df1_combined))

# Extract indicator column
df_indicator <- df2[[1]]
# Remove indicator column from dataframe.
df2 <- df2[,-1]
df2 <- data.frame(t(df2))

```


## ROI Pixel-wise Time-Series Analysis 

```{r, warning=FALSE, timeit = TRUE, eval=FALSE}
# Make a df for collecting pixel-wise time-series statistics. 
# The seven statistic features are :
#1  Average fluorescent intensity of raw series; 
#2  Standard deviation of fluorescent intensity raw series;
#3  Percentage of Eigenvalues 1 ; 
#4  Period 1; 
#5  average Amplitude 1;
#6  Percentage of Eigenvalues 2; 
#7  Period 2; 
#8  average Amplitude 2;
#9  Percentage of Eigenvalues 3; 
#10  Period 3; 
#11  average Amplitude 3;
#12  Percentage of Eigenvalues 4  ; 
#13  Period 4; 
#14  average Amplitude 4

df_stats <- data.frame(t(df2[1:14,]))
names(df_stats) <- c("I_bar","I_std","E1","P1","A1","E2","P2","A2","E3","P3","A3","E4","P4","A4")
# Set all values to 0
df_stats[] <- 0     
# Calculate the average fluorescence intensity (I_bar)
df_stats[,'I_bar'] <- colMeans(df2) 
# Calculate the standard deviation of fluorescence intensity (I_std)
df_stats[,'I_std'] <- apply(df2, MARGIN = 2, FUN=sd)
# Scan through each pixel
for (i in 1:length(df2)) {
  # First step, linear detrend 
  df_pixel_dt <- lm.detrend(df2[i])
  # Second step, Singular Spectrum Analysis (SSA).  
  s <- ssa(df_pixel_dt[["Intensity"]], L = window, plot=FALSE) 
 
  # Calulate singualr vaules in percentage
  SS_percentage <- s$sigma/sum(s$sigma)
  df_stats[i,'E1'] <- SS_percentage[1] + SS_percentage[2]  # Assign the e1 percentage
  df_stats[i,'E2'] <- SS_percentage[3] + SS_percentage[4]  # Assign the e2 percentage
  df_stats[i,'E3'] <- SS_percentage[5] + SS_percentage[6]  # Assign the e3 percentage
  df_stats[i,'E4'] <- SS_percentage[7] + SS_percentage[8]  # Assign the e4 percentage   
  # Extract the left matrix (time-series)  
  recon <- reconstruct(s, groups = list(c(1:2),c(3:4),c(5:6),c(7:8)))
  F1 <- recon$F1
  F2 <- recon$F2
  F3 <- recon$F3
  F4 <- recon$F4
  
  # Calculate average Amplitude of sinusoid time-series
  df_stats[i,'A1'] <- round(mean(abs(F1))*pi/2,digits=2) 
  df_stats[i,'A2'] <- round(mean(abs(F2))*pi/2,digits=2)
  df_stats[i,'A3'] <- round(mean(abs(F3))*pi/2,digits=2)
  df_stats[i,'A4'] <- round(mean(abs(F4))*pi/2,digits=2)

  # Determine period for each principal time-series using ESPRIT (Estimate of Singal Parameters by Rotational Invariance Techniques)
  per_1 <- max(parestimate(s, list(1:2), method = "esprit")$periods)
  per_2 <- max(parestimate(s, list(3:4), method = "esprit")$periods)
  per_3 <- max(parestimate(s, list(5:6), method = "esprit")$periods)
  per_4 <- max(parestimate(s, list(7:8), method = "esprit")$periods)
  
  # Calculate autocorrelation function (ACF) of principal time-series   
  df_stats[i,'P1'] <- sine.screen(F1, per_1, threshold)              # Assign the p1
  df_stats[i,'P2'] <- sine.screen(F2, per_2, threshold)              # Assign the p2
  df_stats[i,'P3'] <- sine.screen(F3, per_3, threshold)              # Assign the p3
  df_stats[i,'P4'] <- sine.screen(F4, per_4, threshold)              # Assign the p4
}

# Convert Header of df1 into X, Y columns of df_stats
rowNames_stats <- rownames(df_stats)
df.xy <- data.frame(t(data.frame(lapply(rowNames_stats, row.xy))))
names(df.xy) <- c("X","Y")
df_stats[["X"]] <- df.xy[["X"]]
df_stats[["Y"]] <- df.xy[["Y"]]

# Retreive the indicator column 
df_stats[['Indicator']] <- df_indicator

# Round Period values to two digits after decimal point
df_stats[['P1']] <- round(df_stats[['P1']], digits=2)
df_stats[['P2']] <- round(df_stats[['P2']], digits=2)
df_stats[['P3']] <- round(df_stats[['P3']], digits=2)
df_stats[['P4']] <- round(df_stats[['P4']], digits=2)

# Determine the characteristic Period for each pixel (maximal period chosen)
df_stats[["Period"]] <- apply(df_stats[,c('P1','P2','P3','P4')], 1, FUN=max)
# Determine corresponding average Amplitude for that maximal Period 
df_stats[["Amplitude"]] <- apply(df_stats, 1, FUN=amplitude_sel)

# Count the number of non-zero terms in each column of df_stats
print("The number of non-zero rows of each column in the statistics dataframe:")
colSums(df_stats[,c(1,2,4,7,10,13,18)] != 0) 

# Show all column names 
#colnames(df_stats)
# Show unique values of period
#print("The unique periodic values obtained from analysis are:")
#sort(unique(df_stats[["Period"]]))

```

## Write and save df_stats into csv file
```{r, timeit = TRUE, eval=FALSE}
write.csv(df_stats[df_stats$Indicator == 0,], file = "BG-1_stats.csv")
write.csv(df_stats[df_stats$Indicator == 1 | df_stats$Indicator == 2,], file = "Cell-9&10_stats.csv")
```

## Load Pre-calculated df_stats dataframe 
```{r, echo = FALSE}
df_stats_cell <- read.csv("Cell-2_stats.csv", row.names = 1)    
df_stats_BG <- read.csv("BG-1_stats.csv", row.names = 1)
df_stats <- rbind(df_stats_cell,df_stats_BG)

# Count the number of non-zero terms in each column of df_stats
print("The number of non-zero rows of each column in the statistics dataframe:")
colSums(df_stats[,c(1,2,4,7,10,13,18)] != 0) 
```


## Image Plot

```{r, warning=FALSE, fig.height = 3, fig.width = 4, fig.align ="center", timeit=TRUE, eval=FALSE}

# Make a copy of df_stats for plotting heatmap image
df_stats_heatmap <- df_stats[df_stats$Indicator == 1 | df_stats$Indicator == 2, ]

# Cell Periodic value and distribution 
matrix_period <- as.matrix.xtabs(xtabs( Period ~ X+Y, data=df_stats_heatmap))/160
matrix_period <- normalize(matrix_period)

# Cut-off background pixel's period value (with Cell outline mask)
df_stats_heatmap[df_stats_heatmap$Indicator == 1,]$Period <- 0

matrix_period_mask <- as.matrix.xtabs(xtabs( Period ~ X+Y, data=df_stats_heatmap))/160
matrix_period_mask <- normalize(matrix_period_mask)

m_period_mask <- abind(matrix_period, matrix_period_mask, along = 3)
img_period_mask <- Image(m_period_mask,colormode = 'Grayscale')
img_period_map_mask <- rgbImage(green=img_period_mask)
display(img_period_map_mask, method= 'raster', all=TRUE)

writeImage(img_period_mask, "cell-4-period-m.jpeg", quality = 100)

# Cell Amplitude value and distribution 
matrix_amp <- as.matrix.xtabs(xtabs( Amplitude ~ X+Y, data=df_stats_heatmap))/80
#matrix_amp <- normalize(matrix_amp)

# Cut-off background pixel's Amplitude value (with Cell outline mask)
df_stats_heatmap[df_stats_heatmap$Indicator == 1,]$Amplitude <- 0

matrix_amp_mask <- as.matrix.xtabs(xtabs( Amplitude ~ X+Y, data=df_stats_heatmap))/80
#matrix_amp_mask <- normalize(matrix_amp_mask)

m_amp_mask <- abind(matrix_amp, matrix_amp_mask, along = 3)
img_amp_mask <- Image(m_amp_mask,colormode = 'Grayscale')
img_amp_map_mask <- rgbImage(blue=img_amp_mask)
display(img_amp_map_mask, method= 'raster', all=TRUE)

writeImage(img_amp_map_mask, "cell-4-amp.jpeg", quality = 100)

# Divide pixels into 4 levels based on period value (frames)
df_stats_heatmap_factor <- df_stats_heatmap
df_stats_heatmap_factor$Period <- cut(df_stats_heatmap_factor$Period, c(0, 8, 24, 48, 72, 96, 120, 180), right = FALSE)
df_stats_heatmap_factor$P1 <- cut(df_stats_heatmap_factor$P1, c(0, 8, 24, 48, 72, 96, 120, 180), right = FALSE)
df_stats_heatmap_factor$P2 <- cut(df_stats_heatmap_factor$P2, c(0, 8, 24, 48, 72, 96, 120, 180), right = FALSE)
df_stats_heatmap_factor$P3 <- cut(df_stats_heatmap_factor$P3, c(0, 8, 24, 48, 72, 96, 120, 180), right = FALSE)
df_stats_heatmap_factor$P4 <- cut(df_stats_heatmap_factor$P4, c(0, 8, 24, 48, 72, 96, 120, 180), right = FALSE)

# Level 1 
df_stats_heatmap_l1 <- df_stats_heatmap
df_stats_heatmap_l1[!(df_stats_heatmap_factor$P1 %in% c("[8,24)","[24,48)")), ]$P1 <- 0
df_stats_heatmap_l1[!(df_stats_heatmap_factor$P2 %in% c("[8,24)","[24,48)")), ]$P2 <- 0
df_stats_heatmap_l1[!(df_stats_heatmap_factor$P3 %in% c("[8,24)","[24,48)")), ]$P3 <- 0
df_stats_heatmap_l1[!(df_stats_heatmap_factor$P4 %in% c("[8,24)","[24,48)")), ]$P4 <- 0
df_stats_heatmap_l1$Period <- apply(df_stats_heatmap_l1[,c('P1','P2','P3','P4')], 1, FUN=max)

# Cut-off background pixel's period value (with Cell outline mask)
df_stats_heatmap_l1[df_stats_heatmap$Indicator == 1,]$Period <- 0

matrix_period_mask_l1 <- as.matrix.xtabs(xtabs( Period ~ X+Y, data=df_stats_heatmap_l1))/160

# Level 2
df_stats_heatmap_l2 <- df_stats_heatmap
df_stats_heatmap_l2[!(df_stats_heatmap_factor$P1 %in% c("[48,72)","[72,96)")), ]$P1 <- 0
df_stats_heatmap_l2[!(df_stats_heatmap_factor$P2 %in% c("[48,72)","[72,96)")), ]$P2 <- 0
df_stats_heatmap_l2[!(df_stats_heatmap_factor$P3 %in% c("[48,72)","[72,96)")), ]$P3 <- 0
df_stats_heatmap_l2[!(df_stats_heatmap_factor$P4 %in% c("[48,72)","[72,96)")), ]$P4 <- 0
df_stats_heatmap_l2$Period <- apply(df_stats_heatmap_l2[,c('P1','P2','P3','P4')], 1, FUN=max)

# Cut-off background pixel's period value (with Cell outline mask)
df_stats_heatmap_l2[df_stats_heatmap$Indicator == 1,]$Period <- 0

matrix_period_mask_l2 <- as.matrix.xtabs(xtabs( Period ~ X+Y, data=df_stats_heatmap_l2))/160

# Level 3
df_stats_heatmap_l3 <- df_stats_heatmap
df_stats_heatmap_l3[!(df_stats_heatmap_factor$P1 %in% c("[96,120)","[120,180)")), ]$P1 <- 0
df_stats_heatmap_l3[!(df_stats_heatmap_factor$P2 %in% c("[96,120)","[120,180)")), ]$P2 <- 0
df_stats_heatmap_l3[!(df_stats_heatmap_factor$P3 %in% c("[96,120)","[120,180)")), ]$P3 <- 0
df_stats_heatmap_l3[!(df_stats_heatmap_factor$P4 %in% c("[96,120)","[120,180)")), ]$P4 <- 0
df_stats_heatmap_l3$Period <- apply(df_stats_heatmap_l3[,c('P1','P2','P3','P4')], 1, FUN=max)

# Cut-off background pixel's period value (with Cell outline mask)
df_stats_heatmap_l3[df_stats_heatmap$Indicator == 1,]$Period <- 0

matrix_period_mask_l3 <- as.matrix.xtabs(xtabs( Period ~ X+Y, data=df_stats_heatmap_l3))/160

#  2 X 2 Display of 3 levels + Combined
m_period_mask_cmd <- abind(matrix_period_mask,matrix_period_mask_l1,matrix_period_mask_l2, matrix_period_mask_l3,along = 3)
img_period_mask_cmd <- Image(m_period_mask_cmd,colormode = 'Grayscale')
img_period_map_mask_cmd <- rgbImage(green=img_period_mask_cmd)
display(img_period_map_mask_cmd, method= 'raster', nx = 2, all=TRUE)

# Cell-1 Sum Fluorescence Intensity Image
matrix_sum <- as.matrix.xtabs(xtabs( I_bar ~ X+Y, data=df_stats_heatmap))
matrix_sum <- normalize(matrix_sum)

# NB Analysis
matrix_BN <- as.matrix.xtabs(xtabs( (I_std)^2/I_bar ~ X+Y, data=df_stats_heatmap))
matrix_BN <- normalize(matrix_BN)

m2_sumNB <- abind(matrix_sum,matrix_BN, along=3)
img_sumNB <- Image(m2_sumNB,colormode='Grayscale')
display(img_sumNB, method = 'raster', all=TRUE)

writeImage(img_sumNB, "cell-4-NB.jpeg", quality = 100)


# Overlay Cell fluorescence summation/NB image and Periodic values 
img <- rgbImage(blue=m2_sumNB, green=m2_sumNB+abind(matrix_period_mask,matrix_period_mask, along=3), red=m2_sumNB)
display(img, method = 'raster', all=TRUE)

# Overlay Cell fluorescence summation/NB image and Amplitude values 
img_2 <- rgbImage(blue=m2_sumNB+abind(matrix_amp_mask,matrix_amp_mask, along=3), green=m2_sumNB, red=m2_sumNB)
display(img_2, method = 'raster', all=TRUE)

#img <- rgbImage(blue=matrix_period_mask_l1*2, green=matrix_period_mask_l2, #red=matrix_period_mask_l3/1.5)
#display(img, method = 'raster', all=TRUE)

# Illustration of ROI and Peripheral
#matrix_ROI <- as.matrix.xtabs(xtabs( (Indicator-1) ~ X+Y, data=df_stats_heatmap))
#matrix_Peripheral <- as.matrix.xtabs(xtabs( -(Indicator-2) ~ X+Y, data=df_stats_heatmap))
#img_ROI <- rgbImage( blue=matrix_ROI*0.4+matrix_sum, #green=matrix_Peripheral*0.4+matrix_sum, red=matrix_sum )
#display(img_ROI, method= 'raster', all=TRUE)

```


## Prepare Dataframe for Plotting the Statistics 

```{r}
# Divide the df_stats into Cellular region and BG region according to indicator
df_stats_Cell <- df_stats[which(df_stats$Indicator == 2),]
df_stats_RI <- df_stats[which(df_stats$Indicator == 1),]
df_stats_BG <- df_stats[which(df_stats$Indicator == 0),]
```

## Plotting Statistics
**Histogram**
```{r, warning=FALSE, fig.height = 4, fig.width = 8,fig.align="center",timeit = TRUE}

df_cmd <- rbind(df_stats_Cell,df_stats_RI, df_stats_BG)

h1 <- ggplot(df_cmd, aes(x=I_bar))  +
  geom_histogram(data = df_stats_Cell, aes(y=..count../sum(..count..)*100), colour="blue", fill = "blue", alpha =0.3, binwidth = 10) +
  geom_histogram(data = df_stats_RI, aes(y=..count../sum(..count..)*100), colour="green", fill = "green", alpha =0.3, binwidth = 10) +
  geom_histogram(data = df_stats_BG, aes(y=..count../sum(..count..)*100), colour="red", fill = "red", alpha =0.3, binwidth = 10) +
  xlab("Average of Fluorescence Intensity") +
  ylab("Percentage (%)")

h2 <- ggplot(df_cmd, aes(x=I_std))  +
  geom_histogram(data = df_stats_Cell, aes(y=..count../sum(..count..)*100), colour="blue", fill = "blue", alpha =0.3, binwidth = 2) +
  geom_histogram(data = df_stats_RI, aes(y=..count../sum(..count..)*100), colour="green", fill = "green", alpha =0.3, binwidth = 2) +
  geom_histogram(data = df_stats_BG, aes(y=..count../sum(..count..)*100), colour="red", fill = "red", alpha =0.3, binwidth = 2) + 
  xlab("Standard Deviation of Fluorescence Intensity") + 
  ylab("Percentage (%)") +
  scale_x_continuous(limits = c(40, 160))

grid.arrange(h1,h2,ncol=2)

```



**Scatter Plot**
```{r,warning=FALSE, fig.height = 7, fig.width = 7, fig.align="center",timeit = TRUE}
  
sp1 <- ggplot(df_cmd, aes(y = I_bar, x = E1*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E1") + 
  ylab("Average of Fluorescence Intensity") 

sp2 <- ggplot(df_cmd, aes(y = I_bar, x = E2*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E2") + 
  ylab("Average of Fluorescence Intensity") 

sp3 <- ggplot(df_cmd, aes(y = I_bar, x = E3*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E3") + 
  ylab("Average of Fluorescence Intensity") 

sp4 <- ggplot(df_cmd, aes(y = I_bar, x = E4*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
   xlab("Percentage (%) of Eigenvalue E4") + 
  ylab("Average of Fluorescence Intensity") 

grid.arrange(sp1,sp2,sp3,sp4,ncol=2)

```

```{r,warning=FALSE, fig.height = 7, fig.width = 7, fig.align="center",timeit = TRUE}
  
sp5 <- ggplot(df_cmd, aes(y = A1, x = E1*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E1") + 
  ylab("Average of Amplitude A1") 

sp6 <- ggplot(df_cmd, aes(y = A2, x = E2*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E2") + 
  ylab("Average of Amplitude A2") 

sp7 <- ggplot(df_cmd, aes(y = A3, x = E3*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.8, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.8, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.8, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E3") + 
  ylab("Average of Amplitude A3") 

sp8 <- ggplot(df_cmd, aes(y = A4, x = E4*100))  +
  geom_point(data = df_stats_BG, color = "red", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_RI, color = "green", size=0.6, alpha =0.3) +
  geom_point(data = df_stats_Cell, color = "blue", size=0.6, alpha =0.3) +
  xlab("Percentage (%) of Eigenvalue E4") + 
  ylab("Average of Amplitude A4") 

grid.arrange(sp5,sp6,sp7,sp8,ncol=2)


```


**Barplot**
```{r, warning=FALSE, fig.height = 4, fig.width = 6, fig.align="center",timeit = TRUE}

ggplot(df_cmd, aes(x=Period*frame_rate/1000))  + scale_x_continuous(limits = c(6, 150)) + 
  geom_histogram(data = df_stats_Cell, aes(y=..count../sum(..count..)*100), colour="blue", fill = "blue", alpha =0.4, binwidth = 2) +
  geom_histogram(data = df_stats_RI, aes(y=..count../sum(..count..)*100), colour="green", fill = "green", alpha =0.4, binwidth = 2) +
  geom_histogram(data = df_stats_BG, aes(y=..count../sum(..count..)*100), colour="red", fill = "red", alpha =0.4, binwidth = 2) +
  facet_grid(Indicator ~ .) +
  xlab("Period (s)") +
  ylab("Percentage (%)")


ggplot(df_cmd[df_cmd$Amplitude!=0,], aes(x=Amplitude))  + scale_x_continuous(limits = c(0, 120)) + 
  geom_histogram(data = df_stats_Cell[df_stats_Cell$Amplitude!=0,], aes(y=..count../sum(..count..)*100), colour="blue", fill = "blue", alpha =0.3, binwidth = 1) +
  geom_histogram(data = df_stats_RI[df_stats_RI$Amplitude!=0,], aes(y=..count../sum(..count..)*100), colour="green", fill = "green", alpha =0.3, binwidth = 1) +
  geom_histogram(data = df_stats_BG[df_stats_BG$Amplitude!=0,], aes(y=..count../sum(..count..)*100), colour="red", fill = "red", alpha =0.3, binwidth = 1) +
  facet_grid(Indicator ~ .) +
  xlab("Average Amplitude") +
  ylab("Percentage (%)")


```



```{r,warning=FALSE, fig.height = 4, fig.width = 10, fig.ncol = 1, fig.align = "center",timeit = TRUE}
# Make a df_stats copy for factors plotting
df_stats_factor <- df_stats[df_stats$Period!=0,]

df_stats_factor$Period <- cut(df_stats_factor$Period*frame_rate/1000, c(6, 20, 40, 60, 80, 100, 120, 150), right = FALSE)

df_stats_factor_Cell <- df_stats_factor[which(df_stats_factor$Indicator == 2),]
df_stats_factor_RI <- df_stats_factor[which(df_stats_factor$Indicator == 1),]
df_stats_factor_BG <- df_stats_factor[which(df_stats_factor$Indicator == 0),]

p5 <- ggplot(df_stats_factor_Cell, aes(y = I_bar, x = Period)) + geom_boxplot(fill='blue',alpha=0.6) + scale_y_continuous(limits = c(1100, 1640)) + 
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Cell ROI") + ylab("Average of Fluorescence Intensity") + xlab("Period (s)") + theme(axis.text.x=element_text(size=6))

 
p7 <- ggplot(df_stats_factor_BG, aes(y = I_bar, x = Period)) + geom_boxplot(fill='red',alpha=0.6) + scale_y_continuous(limits = c(1100, 1640)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Background") + ylab("Average of Fluorescence Intensity") + xlab("Period (s)") + theme(axis.text.x=element_text(size=6))


p6 <- ggplot(df_stats_factor_RI, aes(y = I_bar, x = Period)) + geom_boxplot(fill='green',alpha=0.6) + scale_y_continuous(limits = c(1100, 1640)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Peripheral") + ylab("Average of Fluorescence Intensity") + xlab("Period (s)") + theme(axis.text.x=element_text(size=6))


grid.arrange(p5,p6,p7,ncol=3)

p7 <- ggplot(df_stats_factor_Cell, aes(y = I_std, x = Period)) + geom_boxplot(fill='blue',alpha=0.6)  + scale_y_continuous(limits = c(50, 170)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Cell ROI") + ylab("Standard Deviation of Fluorescence Intensity") + xlab("Period (s)") + theme(axis.text.x=element_text(size=6))


p9 <- ggplot(df_stats_factor_BG, aes(y = I_std, x = Period)) + geom_boxplot(fill='red',alpha=0.6) + scale_y_continuous(limits = c(50, 170)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Background") + ylab("Standard Deviation of Fluorescence Intensity") + xlab("Period (s)") + theme(axis.text.x=element_text(size=6))


p8 <- ggplot(df_stats_factor_RI, aes(y = I_std, x = Period)) + geom_boxplot(fill='green',alpha=0.6) + scale_y_continuous(limits = c(50, 170)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Peripheral") + ylab("Standard Deviation of Fluorescence Intensity") + xlab("Period (s)") + theme(axis.text.x=element_text(size=6))


grid.arrange(p7,p8,p9,ncol=3)

p11 <- ggplot(df_stats_factor_Cell, aes(y = Amplitude, x = Period)) + geom_boxplot(fill='blue',alpha=0.6) + scale_y_continuous(limits = c(10, 100)) + 
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Cell ROI") + ylab("Average Amplitude") + xlab("Period (s)")+ theme(axis.text.x=element_text(size=6))

 
p12 <- ggplot(df_stats_factor_BG, aes(y = Amplitude, x = Period)) + geom_boxplot(fill='red',alpha=0.6) + scale_y_continuous(limits = c(10, 100)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Background") + ylab("Average Amplitude") + xlab("Period (s)")+ theme(axis.text.x=element_text(size=6))


p13 <- ggplot(df_stats_factor_RI, aes(y = Amplitude, x = Period)) + geom_boxplot(fill='green',alpha=0.6) + scale_y_continuous(limits = c(10, 100)) +
    stat_summary(fun.y=mean, geom="point", shape=5, size=2) + ggtitle("Peripheral") + ylab("Average Amplitude") + xlab("Period (s)")+ theme(axis.text.x=element_text(size=6))


grid.arrange(p11,p13,p12,ncol=3)

```

## Checking Individual Pixel

```{r, echo = FALSE, eval=FALSE}

# Input of pixel header
pixel.header <- "X23.Y27"
# Load single time-series from df1
df_pixel <- df2[pixel.header]

# Plot detrend and normalized fluorescent intensity time-series
df_pixel_dt <- data.frame(lm.detrend(df_pixel))
# FFT of the detrend and normalized fluorescent intensity time-series
ssp_dt <- spectrum(df_pixel_dt[["Intensity"]], plot=FALSE)
# Plot the periodgram (up to 64 frames) of the detrend fluorescent intensity time-series
df_ssp_dt <- data.frame(cbind(ssp_dt$spec,round(1/ssp_dt$freq*frame_rate/1000,digits = 1)))
names(df_ssp_dt) <- c('Spec_density','Period')

p3 <- ggplot(df_ssp_dt[2:125,], aes(x=Period, y=Spec_density)) + geom_bar(stat="identity", color="black") + ggtitle("Periodgram of Linear Detrend Fluorescence Signal") + xlab("Period (s)")

p1 <- ts_plot(df_pixel, paste("Pixel", pixel.header, "Fluorescent Intensity Time-Series"))
p2 <- ggplot(df_pixel_dt, aes(x=Time, y=Intensity))  + geom_line() + ggtitle(paste("Pixel", pixel.header, "Linear Detrend Fluorescent Intensity Time Series"))
grid.arrange(p1,p2,p3,ncol=1)

# Conduct SSA
s <- ssa(df_pixel_dt[["Intensity"]], L=window)   

plot(s) # Eigenvalues
plot(s, type = "vectors") # Eigenvectors
plot(s, type = "paired") # Pairs of eigenvectors
plot(wcor(s)) # w-correlation matrix plot

plot(wcor(s, groups = as.list(1:8))) # w-correlation matrix plot

recon <- reconstruct(s, groups = list(c(1:2), c(3:4),c(5:6),c(7:8)))

phase_sine <- function (ts, per){
  t <- seq(0,length(ts)-1)
  df_sine <- data.frame(cbind(ts, sin(2*pi/per*t), cos(2*pi/per*t)))
  names(df_sine) <- c('y', 'sine', 'cos')
  reslm <- lm(y ~  sine + cos , data = df_sine)
  print(summary(reslm))
  
  a2 <- reslm$coefficients[2]
  b2 <- reslm$coefficients[3]
  phase <- as.numeric(atan(-b2/a2)/pi)
  
  df_sine[['y_hat']] <- data.frame(fitted(reslm))

  print(ggplot(df_sine, aes(x = t)) + geom_point(aes(y=y), shape = 1) + geom_line(aes(y=y_hat, color = "red")) + xlab("Time") + ylab("F1") )
  
  return(phase)
}

F1 <- recon$F1
F2 <- recon$F2
F3 <- recon$F3
F4 <- recon$F4
res <- residuals(recon)

F1_amp <- mean(abs(F1))
F2_amp <- mean(abs(F2))
F3_amp <- mean(abs(F3))
F4_amp <- mean(abs(F4))

pF1<- ts_plot(F1,"F1")
pF2<- ts_plot(F2,"F2")
pF3<- ts_plot(F3,"F3")
pF4<- ts_plot(F4,"F4")
pres <- ts_plot(res,"Residue")

grid.arrange(p2,pF1,pF2, ncol=1)
grid.arrange(p2,pF3,pF4, ncol=1)

per1 <- (parestimate(s, list(5:6), method = "esprit")$periods)

per1 <- max(parestimate(s, list(1:2), method = "esprit")$periods)
per2 <- max(parestimate(s, list(3:4), method = "esprit")$periods)
per3 <- max(parestimate(s, list(5:6), method = "esprit")$periods)
per4 <- max(parestimate(s, list(7:8), method = "esprit")$periods)

# Extract the left matrix (time-series)
sine.screen.plot(F1, per1, 0,"F1")
sine.screen.plot(F2, per2, 0,"F2")
sine.screen.plot(F3, per3, 0,"F3")
sine.screen.plot(F4, per4, 0,"F4")


#ssp<-spectrum(F1, plot=FALSE)
#per_fft_1 <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
#ssp<-spectrum(F2, plot=FALSE)
#per_fft_2 <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
#ssp<-spectrum(F3, plot=FALSE)
#per_fft_3 <- 1/ssp$freq[ssp$spec==max(ssp$spec)]
#ssp<-spectrum(F4, plot=FALSE)
#per_fft_4 <- 1/ssp$freq[ssp$spec==max(ssp$spec)]

#sine.screen.plot(F1, per_fft_1, 0,"F1")
#sine.screen.plot(F2, per_fft_2, 0,"F2")
#sine.screen.plot(F3, per_fft_3, 0,"F3")
#sine.screen.plot(F4, per_fft_4, 0,"F4")

```


## Check random pixels with non-zero period values in Cell region

```{r, warning=FALSE, fig.height = 4, fig.width = 7, fig.align="center", timeit=TRUE, eval=FALSE}
set.seed(16)
df_stats_cell <- df_stats[which(df_stats$Indicator == 2),]
df_stats_factor_cell <- df_stats_cell
df_stats_factor_cell$Period <- cut(df_stats_factor_cell$Period, c(0, 8, 24, 48, 72, 96, 120, 160), right = FALSE)

s1 <- df_stats_cell[df_stats_factor_cell$Period %in% c("[0,8)"), ]
s2 <- df_stats_cell[df_stats_factor_cell$Period %in% c("[8,24)","[24,48)"), ]
s3 <- df_stats_cell[df_stats_factor_cell$Period %in% c("[48,72)","[72,96)"), ]
s4 <- df_stats_cell[df_stats_factor_cell$Period %in% c("[96,120)","[120,160)"), ]

s1 <- s1[sample(nrow(s1),1),]
s2 <- s2[sample(nrow(s2),1),]
s3 <- s3[sample(nrow(s3),1),]
s4 <- s4[sample(nrow(s4),1),]

df_stats_cell_sample <- rbind(s1,s2,s3,s4)

for (i in 1:dim(df_stats_cell_sample)[1]) {
  # For pixel with non-zero period value
    header <- rownames(df_stats_cell_sample)[i]
    # Load single time-series from df1
    df_pixel <- df2[header]
    print(paste("The pixel is", header,",", "Its period is", df_stats_cell_sample[i,'Period']*frame_rate/1000))
    # Plot raw fluorescent intensity time-series
    p1 <- ts_plot(df_pixel, paste(header, " Raw Fluorescent Intensity Time-Series"))
    # Plot detrend and normalized fluorescent intensity time-series
    df_pixel_dt <- data.frame(lm.detrend(df_pixel))
    p2 <- ts_plot(df_pixel_dt, paste(header, " Linear Detrend Fluorescent Intensity Time-Series"))
    # Plot spectrum periodgram of linear detrend fluorescent signal (before SSA)
    ssp_dt <- spectrum(df_pixel_dt["Intensity"], plot=FALSE)
    df_ssp_dt <- data.frame(cbind(ssp_dt$spec,round(1/ssp_dt$freq*frame_rate/1000,digits = 1)))
    names(df_ssp_dt) <- c('Spec_density','Period')
    p3 <- ggplot(df_ssp_dt[2:125,], aes(x=Period, y=Spec_density)) + geom_bar(stat="identity", color="black") + ylab("Spectral Density") + ggtitle("Periodgram") + xlab("Period (s)")
  
    grid.arrange(p1,p2,p3,ncol=1)
    
    s <- ssa(df_pixel_dt[["Intensity"]], L=window)     
    plot(wcor(s, groups = as.list(1:8))) # w-correlation matrix plot
    recon <- reconstruct(s, groups = list(c(1:2), c(3:4),c(5:6),c(7:8)))

    F1 <- recon$F1
    F2 <- recon$F2
    F3 <- recon$F3
    F4 <- recon$F4
    
    per1 <- max(parestimate(s, list(1:2), method = "esprit")$periods)
    per2 <- max(parestimate(s, list(3:4), method = "esprit")$periods)
    per3 <- max(parestimate(s, list(5:6), method = "esprit")$periods)
    per4 <- max(parestimate(s, list(7:8), method = "esprit")$periods)

    # Extract the left matrix (time-series)
    sine.screen.plot(F1, per1, threshold,"F1")
    sine.screen.plot(F2, per2, threshold,"F2")
    sine.screen.plot(F3, per3, threshold,"F3")
    sine.screen.plot(F4, per4, threshold,"F4")
}

```







